### 2021-10-04 paper.tex
Adding modifications from JL's version

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{url}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}
\title{
What should I notice? Evaluating the memorability of events for abductive inference.
}
\author{
  \IEEEauthorblockN{
\IEEEauthorrefmark{1}\IEEEauthorrefmark{2} Étienne Houzé,
\IEEEauthorrefmark{1} Jean-Louis Dessalles,
\IEEEauthorrefmark{1} Ada Diaconescu,
\IEEEauthorrefmark{2} David Menga
  }

  \IEEEauthorblockA{\IEEEauthorrefmark{1}
    Télécom Paris, IP Paris, \emph{Palaiseau}, France\\
email: \{first\}.\{second\}@telecom-paris.fr}
  \IEEEauthorblockA{\IEEEauthorrefmark{2} EDF R\&{}D, \emph{Palaiseau}, France\\
email: \{first\}.\{second\}@edf.fr}
}

\maketitle

\begin{abstract}
  When confronted to an unprecedented situation, humans typically show good
  performance in quickly identifying noticeable past events and proposing them
  as possible causal hypotheses. This kind of abductive inference is widely
  overlooked in modern AI approaches which rely on massive datasets to learn
  associative patterns. Our proposal is to formalize and compute a ``memorability''
score over a set of events recorded from a cyber-physical system.
This score can then be used to select relevant information to be
remembered and to propose causal hypotheses in unusual situations on demand.
The approach is meant to be complementary to more traditional
learning-focused abduction techniques. We provide theoretical ground for our approach
based on Algorithmic Information
Theory. We also provide an implementation example, a
smart-home scenario, to show how our approach could translate in practice.
\end{abstract}

\begin{IEEEkeywords}
Kolmogorov complexity, Algorithmic Information Theory, Simplicity, Abduction, Surprise
\end{IEEEkeywords}

\section{Introduction}

% As the number of connected devices and sensors grows, so does the amount of
% produced data. As a consequence, the ability to filter out a few events from past recordings
% as ``memorable'' becomes prime. While this task is often overlooked by many
% statistical approaches, which instead benefit from large quantity of
% information, it is part of the learning process for humans
% \cite{dehaene_how_2020}. Faced with new situations, humans are prone to infer
% previously seen unusual events as possible causes, and carry out further testing
% to assess the existence of a causal relation.

As a user has just turned on the TV in her all-equipped living-room, the
lights dim and the window blinds go down. Intrigued by this behavior, she
quickly infers that both light dimming and blind closing occurred as a
consequence of the TV being turned on. How did she come to this conclusion?
By performing \emph{abductive inference} \cite{magnani_abduction_2011}. This mental operation
is a key element of humans' ability to understand the world: from the
observed consequences, they infer the possible causes.

In this example, there are mainly three possibilities to come to the
conclusion. (1) If the user knows how the smart living-room system works, if she
knows the underlying rules or parameters, she may use this causal
knowledge to perform abduction. (2) If she has no knowledge
about the system but made several observation of the same behavior, she may examine past
correlations and figure out that turning on the TV set often leads the blinds to
close and the lights to dim. (3) If there are no previous occurrences of the event (e.g. it is the first time she turns on the TV
in the living-room), she may still be able to
suspect that the TV is a possible cause for the observed event, just because it appears to her as a memorable recent event (as it is its first occurrence). This example
suggests that human beings are able to use distinct methods to perform abductive tasks
and infer new knowledge. While the first two mechanisms can be
automated using knowledge bases and statistical methods, ``memorability-based'' abduction needs to be investigated in more depth.

% For instance, it is
% possible, for humans without any prior knowledge about
% physics or electrical engineering, to infer a memorable thunderstorm as a
% possible cause for a general black-out just by using the memorability of
% the former. While this kind of abductive inference (i.e. inference to the cause
% of an observation \cite{magnani_abduction_2011}) can yield many
% false-positives, it can be considered in situations where the lack of knowledge
% from previous occurrences penalizes other approaches.

% The difficulty of this approach to reasoning is, however, to quickly identify
% the relevant candidate hypotheses. In this regards, we may use a basic
% instinct: without prior knowledge, the most relevant hypothesis might simply be
% the most memorable recent event. However this definition is highly subjective,
% and does not seem to fit well into the canvas of computing and AI.

Defining a score of memorability is not straightforward. First, events can be of different nature, and not directly
comparable. Even for restricted systems such as smart homes, noticeable events
range from device removal to presence detection or unusually high
temperatures. Even for comparable events, the problem is to weigh different characteristics: is a record-high temperature 47 days
ago more memorable than the small deviation recorded just 3 minutes ago? To
our knowledge, no current system proposes to combine various
event types from different devices to compute a unified metrics of
``memorability''.

% The difficulty of doing so comes from the wide variety of events, parameters and
% characteristics to take into account. How can one assess if a small recent
% perturbation is more important than a larger but older one? In addition, one
% needs a common metrics to evaluate the relative importance of data coming from
% various devices. In the case of a smart home, for instance, devices range from
% presence sensors to humidity and temperature sensors, power meters, etc. To this
% day, no current system proposes to analyze all events coming from such various
% devices and come to a unified metrics of ``memorability''.

% To tackle this challenge, we use as a starting point the observation that while
% all events, regardless of their characteristics or nature, can be uniquely
% described using a combination of qualifiers, the most memorable ones are likely
% to require less words to be described. Think, for instance, of ``last year's
% hottest day'' and ``the 182th day of 7 years ago''. By evaluating the lenght of
% each description, taking into account both the complexity of concept words (a
% time ranking, a temperature ranking), and the arguments used in both
% descriptions (the hottest, the 182th, 7), we can assign each event a complexity
% score. Memorable events, as they stand out, would then differ from their
% neighbors by being much simpler (or more complex).

To address this issue, we started from the following intuition: while all
events, regardless of their characteristics or nature, can be uniquely
described using a combination of quantitative or qualitative qualifiers, the most memorable ones are
likely to require less words to be described. ``Last year's hottest day'' is simpler than ``the 182\textsuperscript{nd} day 7 years ago''. How can we quantify this relative simplicity? We could evaluate the complexity of each
description, taking into account both the complexity of the concept words (a date
of occurrence, a temperature ranking), and of the arguments (1st hottest,
182 and 7). This approach, The resulting values define the  \emph{description complexity}
of events. Our intuition is that memorable events require simpler and less numerous qualifiers to be
unambiguously described than unremarkable ones.

% A possible approach would be to rely on the naming complexity of events:
% memorable events are more likely to be shorter to be named than boring usual
% events. Think, for instance, of how ``last year's hottest day'' description
% appears much simpler than ``the 182th day of 7 years
% ago''\cite{robles_applications_2010}. Our idea is to rely on this metrics to
% estimate the description complexity of events. Then, by comparing the
% description complexity of a given event $e$ to the average complexity of similar
% events, we induce a measure of unexpectedness that can be used to assess the
% most memorable events.

\begin{figure}[ht]
  \centering
\includegraphics[width=\linewidth]{figures/filters}
\caption{Retrieving an event through successive predicative
filters. From the base memory (yellow), successive filters select events satisfying the associated predicate (grey arrows). For example,
filter $f_1$ selects events from last year, i.e. which satisfy the
predicate $\mathtt{date}(event, 1\_year)$. In this case, successively
applying filters $f_1$, $f_2$ and $f_3$ yields a unique event $e_h$, last
year's hottest day. The complexity of this event can then be upper-bounded
by the complexity of the three filters, since they give an unambiguous way
to describe the event within the memory.}
  \label{fig:filters}
\end{figure}

For machines to implement and compute description complexity, we need a formal framework and computation methods that are in line with human intuition.  Algorithmic Information Theory (AIT) appears to be such a framework, as it is consistent with the human perception of complexity\cite{li_introduction_2008,dessalles2011coincidences,delahaye_numerical_2012}. Our method is as follows: we
consider events as being elements stored in what we call \emph{base memory}. To
reproduce the language features applicable to events, we use \emph
{predicates}, i.e. functions assigning a boolean value to events. For instance, the predicate $\mathtt{date}
  (\cdot, 1\_year)$ is \emph{true} of events that occurred last year. Selecting
all events, from the memory, that satisfy a given predicate corresponds to
a \emph{filter} operation. It generates another memory that is a subset of the previous one.
The filtering operation can then be repeated, selecting
fewer events at each iteration, until a singleton memory is reached. This means that the sequence
of predicates could unambiguously \emph{retrieve} the unique remaining event.
The description complexity of this event can thus be upper-bounded by the
number of bits required to describe the filters used in the retrieval
process. Figure \ref{fig:filters} illustrates this process for the event: ``last year's hottest day''.

% To programmatically achieve this computation, we rely on principles from
% Algorithmic Information Theory, which provides formal definitions of complexity
% and simplicity of objects, which appear to be consistent with human perception
% \cite{li_introduction_2008, dessalles2011coincidences, delahaye_numerical_2012}.
%  Figure \ref{fig:filters} shows our solution applied to the example description
%  ``last year's hottest day''. We assume that
% a memory $\mathcal{M}$ is a unordered set of various distinct objects, upon
% which can be applied boolean predicates $\pi_{k}$ to describe them. These predicates thus
% provide filters $f_{\pi, k}$ that can be applied to the memory, selecting all events
% satisfying predicate $\pi_{k}$. By applying successive
% filters, we narrow down the size of the memory, until a single event remains.
% All the successive filters used to come to this singleton memory constitutes a
% \emph{retrieval path}, whose length, in bits, gives an upper bound of the
% description complexity of the retrieved event.

We will first briefly introduce
some basic notions of Algorithmic Information Theory. Then, in
section~\ref{sec:computing}, we show how we compute the
memorability of past events. We illustrate our approach in~
(section \ref{sec:example}) by describing an implementation
in a smart home simulation context. In section~\ref{sec:related} we review
a few studies that use complexity theory to generate explanations for the smart home
and we discuss how our work can be linked with them. We eventually explore potential
extensions of our work in section \ref{sec:future}

\section{Theoretical Background}
\label{sec:theory}
Kolmogorov complexity formally quantifies the amount of information required
for the computation of a finite binary string\footnote{Though the definition
holds for some infinite binary strings (think of the representation of the
decimals of $\pi$), we restrict ourselves here to finite strings.} (or
any object represented by a finite binary
string)\cite{kolmogorov_three_1965,li_introduction_2008}. The complexity $K(s)$ of a (finite) binary string is the length in bits of the shortest program $p$
which, if given as input to a universal Turing Machine $U$, outputs $s$.
\begin{equation}
  \label{eq:kolmo_def}
  K_{U}(s) = \min_{p}\left\{l(p)|U(p)=s\right\}
\end{equation}

The first notable property of this definition is its universality: while the
choice of the Turing machine $U$ used for the computations appears in the
definition of eq. \ref{eq:kolmo_def}, all results hold, up to an additional
constant, if we change the machine. Think how any Turing-complete programming language can
be turned into any other language, using an interpreter or a compiler program. Since any
Turing machine $U'$ can be emulated by $U$ from a
finite program $p_{U}$, we have the following inequality:
\begin{equation}
  \label{eq:inequality_univ}
K_{U'}(s) \le l(p_{u}) + K_{U}(s)
\end{equation}

From this first result, we can then define complexity $K(s)$, based on the choice of a reference Turing machine, such that, for any other
machine $U$ taken from the set $\text{TM}$ of Turing machines:
\begin{equation}
  \forall U\in\text{TM}, \forall s, |K(s) - K_{u}(s)| \le C_{U}
\end{equation}
where the additional constant $C_{U}$ does not depend on $s$.

Note that the notion of Kolmogorov complexity involves no requirement on the execution time of
programs, only their length in bits matters for the computation of
complexity. Though Kolmogorov complexity can be shown to be
incomputable\cite{li_introduction_2008},
%The sketch of the proof is the
%following: since the definition of complexity from Equation. \ref
%{eq:kolmo_def} requires to find the shortest of all programs on a given Turing
%machine outputting $s$, it requires to run all programs up to a given length
%and compare their result to $s$. However, being able to do so means that we
%would be able to tell whether these programs terminate. Since the termination
%of programs is famously undecidable, by extension, the computation of
%Kolmogorov complexity is impossible.
it can be approximated with
upper bounds by exhibiting a program outputting $s$.

Interestingly, Kolmogorov complexity matches the intuitive
notion and perception of complexity from a human standpoint. For instance, the
complexity of short binary strings evaluated in \cite{delahaye_numerical_2012}
shows similar results to human perception of complex strings and patterns. More
recently, \cite{murena_solving_2020} used Kolmogorov complexity to solve
analogies and showed results close to human expectations.

The bridge between Algorithmic Information Theory (AIT) and human perception of
complexity can be pushed farther thanks to the notions of simplicity and unexpectedness,
which are sometimes regarded of uttermost importance in cognitive science\cite
{chater_simplicity_2003}.
\cite{dessalles2011coincidences} proposes a formal definition of the
unexpectedness $U(e)$ of an event, as the difference between an a-priori
expected causal complexity $K_{w}(e)$ and the actual observed complexity $K
  (e)$.
\begin{equation}
\label{eq:unexpected} U(e) = K_{w}(e) - K(e)
\end{equation}

This result comes from the understanding that, while Kolmogorov complexity is
ideally computed using a Turing machine, it can be used as a proxy for modeling
information processing in the human brain, and thus can be used to define a notion of
simplicity or complexity of events.

Definition \ref{eq:unexpected} allows to model phenomena
such as coincidences: imagine that you happen to run into someone in a park.
If this person has no particular link to you, the event will be quite
trivial: the complexity of describing this person will be equivalent to
distinguishing it from the global population, which is also roughly
equivalent to the (causal) complexity of describing the circumstances having brought this person to be in
that park as the same time as you. On the other hand, if you run into your best friend in
a park, as the complexity of describing your best friend is significantly
lower, the description complexity $K(e)$ drops while the causal complexity
$K_W (e)$ remains unchanged. This is why this latter  event appears unexpected.

Using these insights from AIT, we define the memorability $M(e)$ of an event as the absolute difference between the description complexity $K_d(e)$ of an event and its expected description complexity $K_{exp}(e)$:
\begin{equation}
  \label{eq:memorability}
  M(e) = |K_{exp}(e) - K_d(e)|
\end{equation}
Contrary to the definition of unexpectedness from Equation \ref{eq:unexpected}, we use an absolute value: we do so to acknowledge the fact that events more complex than expected can be memorable as well\footnote{In the original paper \cite{dessalles2011coincidences}, exceptionally complex events are described by considering complexity itself as a way to describe the event: see ``the Pisa Tower effect''\cite{dessalles_pisa_nodate}}. In the next section, we define computational approximations for the description complexity $K_d$ and the expected complexity $K_{exp}$ of events.

\section{Computing the memorability of events}
\label{sec:computing}
\subsection{Retrieving an event}

%Computing the memorability of events begins with the introduction of formal definitions of events, how they are stored, how we can describe them and how these descriptions can be used to retrieve them.
We define \emph{events} as data points augmented with a \emph{label} indicating their nature (temperature event, failure event, addition/removal of a device) and a timestamp of occurrence. Formally:
\begin{equation}
  \label{eq:event}
  e = (l, t,\mathcal{D})
\end{equation}
where $l$ is the label, $t$ the timestamp and $\mathcal{D}$ a multi-dimensional data point representing the various characteristics of $e$: its duration, the maximum temperature reached, the sensor name, its position, etc. Labels can also be considered as classes of events, of which each event is a particular instance.

To model how humans are able to describe events by using qualifiers, we use \emph{predicates}: Boolean functions operating on events and, possibly, additional parameters: $\pi(e, a_1, a_2, \dots, a_n) \mapsto \{O,1\}$ is a predicate of arity $n$ operating on event $e$. In the rest of this paper, we will prefer the equivalent notation $\pi(e, k) \mapsto \{0,1\}$, where $k$ is a binary string encoding the sequence of arguments $a_1, \dots, a_n$. Using this notation, the predicate $\pi$ becomes a boolean function operating on $\mathbf{E} \times \{0,1\}^*$:
\begin{equation}
  \label{eq:predicate}
  \pi : \begin{cases}
\mathcal{M}\times \{0,1\}^{*} &\mapsto \{0,1\} \\
(e, k) &\mapsto \pi_{k}(e)
\end{cases}
\end{equation}
As an example of predicate, consider $\pi = \mathtt{year}$ and $k$ a string encoding the number $1$, thus
constructing the predicate $\mathtt{year(}e, 1\mathtt{)}$, which tells whether
the event $e$ occurred $1$ year ago.

As events occur, they are stored in the  \emph{base memory} $M_0$. As they are not directly comparable, the memory $M_0$ can be considered as having the structure of an unordered set. We denote by $\mathcal{M}$ the subsets of $M_0$. By extension, elements of $\mathcal{M}$, i.e. subsets of $M_0$, are also called \emph{memories}.

By applying a given predicate $\pi$ to all events contained in a memory $M \subseteq M_0$, and selecting only events satisfying $\pi$, one gets another memory $M_1 \subseteq M \subseteq M_0$. We call this operation a \emph{filter}:
\begin{equation}
  \label{eq:filter}
f_{\pi, k}: \begin{cases}
\mathcal{M} & \mapsto \mathcal{M}             \\
M           & \mapsto \{e \in M | \pi(e) \}
\end{cases}
\end{equation}
For instance, using the same $\pi = \mathtt{year}$ and $k=1$ as above, we can
build the filter $f_{\pi, k} = \mathtt{last\_{}year}$, which selects all events
that occurred last year.

As the output of a filter applied to a memory $M$ is another memory
object $M' \subseteq M$, we can compose filter
functions. A sequence of such filters is called a \emph{retrieval path}
\begin{equation}
\label{eq:ret_def}
p = (f_{\pi_{1}, k_{1}}, \dots, f_{\pi_{n}, k_{n}})
\end{equation}
By definition
$p(M) = f_{\pi_{n}, k_{n}}(\dots(f_{\pi_{1}, k_{1}}(M)))$.
In case the result of the operation $p(M)$ contains a single element
$e$, we say that the path $p$ \emph{retrieves} the element $e$ from $M$, and write
$p(M) = e$. In the example shown in Figure \ref{fig:filters}, the three filters $f_1, f_2, f_3$ form a retrieval path retrieving the event ``last year's hottest day'' from the base memory $M_0$.

\subsection{Description complexity of events}

As presented in sec. \ref{sec:theory}, we are interested in computing an approximation of the description complexity of an event $e$. From the above definitions, if there is a path $p$ retrieving $e$ from the base memory $M_0$, i.e. $p(M_0) = e$, this path provides a possible unambiguous description for $e$. We therefore define the description complexity of $e$ as the minimum complexity of a path $p$ retrieving $e$ from the base memory $M_0$.

\begin{equation}
  \label{eq:k_desc}
  K_d(e) = \min_{p \in P_\infty} \{L(p) | p(M_0) = e\}
\end{equation}

where the bit-length $L(p)$ of a retrieval path is defined as the number of bits of a string encoding the path. If we limit ourselves to prefix-free strings encoding predicates and arguments, the total bit length is given by:
\begin{align}
  \label{eq:bit_lenght_p}
  L(p) & = L((f_{\pi_1,k_1}, \dots, f_{\pi_n, k_n}))     \\
       & = L(\pi_1) + L(k_1) + \dots + L(\pi_n) + L(k_n)
\end{align}

By considering only a finite number of possible predicates $\pi$ and arguments $k$, and a maximum path length, we can construct a finite set $P$ of possible retrieval paths. By limiting the search over this set, we get an upper bound of description complexity, and use this upper bound as an approximation:

\begin{equation}
\label{eq:approx_k_desc}
K_d(e) \leq \min_{p \in P \land p(M_0) = e} L(p) = \min_{p \in P \land p(M_0)=e} \sum_{f_{\pi, k} \in p} L(\pi) + L(k)
\end{equation}

\begin{algorithm}
  $\mathtt{current_{explore}} \leftarrow [(\mathcal{M}, 0)]$ \;
  $\mathtt{future_{explore} \leftarrow} [\;]$ \;
  $\mathtt{pass} \leftarrow 0$ \;
  $\mathtt{K(e)} \leftarrow +\infty$ \;
  \While{$\mathtt{current_{explore}} \neq [\;]$ \textbf{and} $\mathtt{pass} < \mathtt{max\_pass}$}{
\For{$(\mathtt{M_{prev}}, \mathtt{K_{prev}}) \in \mathtt{current_{explore}}$}{
\For{$\mathtt{\pi \in \mathcal{P}}$}{
\For{$\mathtt{k \in \{0,1\}^{*}}$}{
$\mathtt{K_{current} \leftarrow l(\pi) + l(k) + K_{prev}}$ \;
\If{$\mathtt{K_{current}} > \mathtt{max_{complex}}$}{
  \textbf{break} \;
}
$\mathtt{M'} \leftarrow f_{\pi,k}(\mathtt{M_{prev}})$ \;
\eIf{$\mathtt{M'} = \mathtt{\{e\}}$}{
  $\mathtt{K(e) \leftarrow \min(K(e), K_{current})}$ \;
}{
  $\mathtt{future_{explore}.append((M', K_{current}))}$\;
}
}
}
}
$\mathtt{current_{explore}} \leftarrow \mathtt{future_{explore}}$ \;
$\mathtt{future_{explore}} \leftarrow [\;]$ \;
$\texttt{pass} \leftarrow \mathtt{pass} + 1$\;
  }
  \caption{Iterative computation of the approximate complexity}
  \label{alg:complex_iter}
\end{algorithm}

The approximation of description complexity from Equation \ref{eq:approx_k_desc} allows for
a direct implementation, which is shown in Algorithm \ref{alg:complex_iter}. This
algorithm operates iteratively: starting from the base memory $M_0$
(line 1), we apply all possible predicate concepts $\pi$ from a given finite set
$\Pi$ and programs $k$ (lines 6-7), up to a given length $\mathtt
{max\_len}$ bits, and apply them: $M' = f_{\pi, k}(M)$ (line 12). We then store
the pairs $(M', \mathtt{len(}\pi, k\mathtt{)})$ in an array $\mathtt{future_
{explore}}$. At the end of the iteration, the results of the filters become the
memories which will be explored during the next iteration(lines 21--23). Each
pass thus explore retrieval paths of increasing length. When a singleton memory
is reached, the complexity of its unique element is upper-bounded with the
length of the corresponding retrieval path (line 14).

\subsection{Computing Memorability}
As stated in Equation \ref{eq:memorability}, we define memorability $M(e)$ as the absolute difference between the description complexity of an event and its expected value. As we've just defined $K_d(e)$ and provided an approximation in Equation \ref{eq:approx_k_desc}, we now focus on defining the \emph{expected} description complexity of an event, $K_{exp}(e)$ that appears in Equation \ref{eq:memorability}

$K_{exp}(e)$ evaluates the complexity the user, or the system, would expect the occurrence of event $e$ to have, based on their previous knowledge. In our framework, this prior knowledge consists of the base memory $M_0$. The expected complexity of the event $e$ can be computed with a simple first-order approximation, i.e. estimating the average complexity of ``similar events'' over the base memory $M_0$.

Still, there is a difficulty in defining what
should be considered \emph{similar} events. Given that we deal with non comparable events, we may define the notion of similarity by referring once again to \emph{predicates}.For a given event $e$ and a given predicate $\pi_k$, we define a $\pi_k$-neighborhood of $e$ as the set $N_{\pi, k}(e)$ of all other events satisfying $\pi_k$.

\begin{equation}
\label{eq:similar}
N_{\pi, k}(e) = \{e'\in M_0, \pi_k(e') \wedge e' \neq e\}
\end{equation}

Now, when considering, for all possible predicates $\pi_k$, the corresponding
neighborhoods $N_{\pi, k}(e)$, with the convention that $N_{\pi, k}(e) = \emptyset$
if $e$ does not satisfy $\pi_k$, we can compute an average expected complexity for $e$:

\begin{equation}
\label{eq:expected}
K_{exp}(e) = \frac{
\sum_{\pi, k} \sum_{e' \in N_{\pi, k}(e)} K_d(e')
}{
\sum_{\pi, k} |N_{\pi, k}(e)|
}
\end{equation}

This definition is consistent with the intuitive idea that more similar events should weigh more in the computation. Indeed, if $e'$
is very similar to $e$, it will appear in many neighborhoods, since it
satisfies most of the predicates that $e$ satisfies. Therefore, it will
be present in more terms in Equation \ref{eq:expected}, and will weigh more in
the final result.

\subsection{From memorability to abduction}
% On peut étendre la définition utilisée pour la surprise en faisant intervenir
% les prédicats'' `sympathiques` par rapport à ce que l'on questionne.
Abductive inference builds on the computation
of the memorability score. \emph{Knowing} that we want to find a cause $c$
for an observed effect $e$, we try to find the most remarkable event in past
memory that is related to $e$. While our ``memorability'' score identifies remarkable past events, it does not take into account their relatedness to $e$.

The knowledge attached to the occurrence of $e$ can be integrated into the description complexity definition by using conditional complexity $K_d(c | e)$: The information contained in $e$ is considered as given, and therefore ``free'' in terms of complexity. For instance, when looking for a cause for
an anomaly in the living-room, other anomalies occurring in the same
living-room will be simpler, as the location ``living-room'' is already known
from the observation of the current anomaly.

Formally, we now consider only predicates where knowledge of the effect event $e$  is appended to all programs $k$:
$\pi_{k::e}(c)$, where $::$ is the append operation. The set of paths obtained with such predicates is noted $P^\infty_e$. This append operation is free in terms of bit-length in the computation of complexity, since the effect event $e$ is an input of the problem. Therefore, we have $L'(\pi_{k::e}) = L(\pi_k) = L(\pi) + L(k)$. We get a definition for the conditional description complexity:

\begin{align}
  \label{eq:abd_k}
  K_d(c | e) & = \min_{p \in P^\infty_e} \{L'(p), \quad p(M_0)=c \}                                                \\
             & = \min_{p \in P^\infty_e} \left\{\sum_{f_{\pi, k::e} \in p} L(\pi) + L(k), \quad p(M_0) = c\right\}
\end{align}

This new conditional description complexity translates the additional information provided to the
system when answering a user's request. It can then be averaged over similar events to compute the expected conditional description complexity, $K_{exp}(c|e)$. From this, we come to the definition of the conditional memorability, which measures how memorable an event $c$ turns out to be in the context of the occurrence of another event $e$:

\begin{equation}
\label{eq:cond_mem}
M(c|e) = |K_{exp}(c|e) - K_d(c|e)|
\end{equation}

Conditional memorability encapsulates the idea
presented as the motivation of this paper: when confronted with a surprising
situation, and in the absence of any other source of information, events that appear more memorable than others, with regards to the target event will be selected as potential causes. As such, our conditional memorability score provides a ranking that can be used for abductive inference.

\textbf{TODO:} Mettre un passage ici pour dire que ça résoud le dilemne de la récence présenté dans l'intro

\section{Implementation Example}
\label{sec:example}
\subsection{Setup}
We considere here an experimental smart home setup. This choice of
configuration is motivated by the challenges posed by smart homes for abductive
inference: i) as the number of connected devices increases, more events are
recorded, making the detection of memorable events more important; ii) smart
homes are prone to experience atypical situations, highly dependent on the context,
for which pre-established relations might fail to find good abduction
candidates. Our choice was also motivated by
the existence of previous work involving smart home simulations capable of quickly generating data from which we could extract events and test our methods.

% In order to provide an acceptable setup to our experiments, we used the scenario
% of smart homes. This kind of scenario is a prime example of situations where
% innovative methods of abduction can prove useful, for various reasons: i) as the
% number of sensors grows with the number of equipped devices in the house, not
% all recorded events are useful and should be remembered over long period of
% time; ii) atypical situations, which are the ones where abduction is most likely
% to be used (to explain situations to the user, or to solve a conflict), are also
% where the lack of past data makes knowledge acquisition hard.

To carry out the simulation, we built custom modules into the existing iCasa
smart home simulator\cite{lalanda_self-aware_2017}. ICasa
offers a simulation of autonomic systems that can handle internal communications,
the possible insertion of new components at runtime, or the deletion or modification of existing
components. We used a basic scenario consisting of a house with four rooms,
a single user, and an outdoor zone. All four rooms are equipped
with a temperature controller system that monitors and controls heaters (fig.
\ref{fig:view}).

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{figures/simulator}
  \caption{View of the simulator's web interface provided by iCasa. The four
    rooms are visible, with their equipment and the user.}
  \label{fig:view}
\end{figure}

Based on this, we implemented a scenario spanning over 420 days, and
comprising a daily cycle of outdoor weather (temperature and sunlight) fluctuations, as well
as user's movements. All these daily changes create non-noticeable events,
serving as a background for our experiments. To produce outstanding
events, we randomly generated about twenty events, spanning over the whole
duration of the simulation, of different kinds:

\begin{itemize}
\item Unusual weather: the outdoor conditions are set to unusually high or
        low temperatures.
\item Heater failures: heater may break down, making them turn off regardless
        of the command they receive.
\item User's absence: the user goes out of the building for an extended
        period of time.
\item Device removal/addition: a device is removed, or another one is added
        to the system.
\end{itemize}

\begin{figure}[ht]
  \includegraphics[width=\linewidth]{figures/ts_example}
  \caption{Time series data from the simulation: outdoor temperature (blue) and
controller temperature of a room (orange). On different occasion, the
    controlled temperature deviates from its setting (21°C). We will evaluate if
our system finds these deviations as memorable, and if it can turn them into relevant causal
hypotheses for current anomalies.}
  \label{fig:ts_example}
\end{figure}

The values observed from all devices and zones was sampled throughout
the simulation run. The resulting data (figure
\ref{fig:ts_example}) was then used as a basis for our experiments.

\subsection{Implementing the complexity computation}

For the implementation of our method, we first needed to identify and
characterize events from the time series data generated by the iCasa simulation.
Since the selection of events is not the focus point of our present work (see sec.
\ref{sec:related}), we perform event detection merely based on threshold and pre-computed conditions.

This set of events constitutes the basis of the initial memory $M_0$ used
for computations.
Then we implemented a small number of predicate concepts:
\begin{itemize}
\item $\mathtt{label(e, k)}$: whether the event $e$ has the label ranked
$k^{th}$ in the label list.
\item $\mathtt{rank(e, r, a)}$: whether the event $e$ has the rank $r$
along axis $a$.
\item $\mathtt{day(e, k)}$: whether the event $e$ occurred $k$ days ago.
\item $\mathtt{month(e, k)}$: whether the event $e$ occurred $k$ months
        ago.
\item $\mathtt{location(e, k)}$: whether the event $e$ occurred in zone
        $k$.
\end{itemize}

With a straightforward implementation of memory, predicates and filters, we could run alg. \ref{alg:complex_iter} worked, though, as expected, it took too
long to be usable in realistic scenarios with hundreds or thousands of events to
consider. In order to facilitate and speed up computations, we also implemented
the following improvements:
\begin{itemize}
  \item The memory object was augmented with various built-in rankings, allowing
for faster operations during future filtering. For instance, since the memory
object keeps a mapping from timestamps to events one can perform a quick
filtering by date without having to loop over all stored element. This convenient mapping,
however, is not directly used to retrieve events by their date of occurrence, so as to
preserve the theoretical model of memory as an unordered set, as presented in
section \ref{sec:computing}.

  \item Each of these predicates holds the property that, in addition to
\texttt{True} and \texttt{False}, they can return another value,
\texttt{None}, which is theoretically treated as \texttt{False} but carries
the additional information that this predicate concept will also be false for
any other element of the memory for any subsequent program $k$. This allows to
effectively break the innermost loop in alg. \ref{alg:complex_iter}.

\item Some of the filters, for instance the date and rank filters, were
hard-written. Events can be selected from these pre-computed mappings over the memory objects
rather than by testing a predicate over all memory elements.
\end{itemize}

\subsection{Results}

\subsubsection{Memorability}

The results of the description complexity evaluation, for the described setup,
are shown in fig. \ref{fig:computed_cplx}. The entire computation, over 4
iterations (meaning that retrieval paths contained at most 4 filters), took
around 30 seconds on a commercial laptop with an i7-8700u CPU.

The blue line emerges from a set of ``usual'' events, whose complexity varies as the logarithm of
the elapsed time since their occurrence. It corresponds to events
for which the best retrieval path consists of a time description (e.g. ``2
months and 12 days ago''). On the other hand, some events stand out in terms of
complexity: some appear simpler, as they can be distinguished by using their
rank along some axis (``the hottest day'', ``the second longest user's
absence''), or the rare occurrence of their kind (``the only fault on the heater'').

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{figures/complexities_computed}
\caption{Computed description complexities of events with retrieval paths of
    length at most 4. Events of type ``day'' (blue), ``hot'' (green), ``cold''
    (red), ``device removal''(cyan) are shown.}
  \label{fig:computed_cplx}
\end{figure}

The ``memorability score'' is shown in fig. \ref{fig:result1}.
% highlights these aforementioned events from the main ``usual'' sequence. Since
the computation of this measure treats unusually complex or unusually simple events the
same way (from the absolute value operation in Equation \ref{eq:unexpected}), we
observe that some events are memorable due to their context only. For instance,
a temperature anomaly occurring simultaneously to many other anomalies is
noticeable, as it is costlier than expected to distinguish it from its neighbors.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{figures/complexities_surprises}
\caption{Memorability score for events in memory}
  \label{fig:result1}
\end{figure}

Given that we generated the data used for this experiment, it is possible to
flag all perturbation events from the usual daily events and evaluate how a
detection based on ``memorability'' score would perform in distinguishing these
events. The result is presented as a ROC curve in fig. \ref{fig:roc}.

\textbf{TOOD:} Il faudrait commenter la courbe
\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{./figures/roc}
  \caption{Experimental ROC curve (True Positive Rate against False Positive
Rate) for a classifier based on our memorability score. Measures consider 23
manually flagged events as memorable (events added to the normal background
as described in Section~\ref{sec:example}.)}
  \label{fig:roc}
\end{figure}
\subsubsection{Abduction}

As an illustration of the abductive inference made possible our memorability
score, we used as a target anomaly the temperature drops visible in the raw
time series data from our scenario (fig \ref{fig:ts_example}).

\textbf{TODO: the experiment still needs to be done: some additional data is required here (creating the scenario and finding causes.)}

\subsubsection{Discussion}
\label{ssub:discussion}

\section{Related Works}
\label{sec:related}
\textbf{TODO: this section is still WIP: I've added most references and short comments, but I still need to write the story linking them.}
Our work is intended to be integrated into larger-scale frameworks to monitor
and detect events in complex environment such as smart homes. In these works, the
approach to smart homes is often regarded as self-organizing systems \cite{kramer_rigorous_2009,kounev_notion_2017}. As such, they present capacities of adaptation to new goals,
new components, new environment. A commonly used approach is the principle of
autonomic system, which minimizes user's intervention for management of the
system \cite{kounev_notion_2017,kephart_vision_2003}.

We want to inscribe our work among other classical concurrent approaches to
abduction. In situations where more data is available, we could for instance rely
on correlation or causal inference from known relations \cite{peters_elements_2017,fadiga_or_2021}.
Previous relations between inference and complexity have been studied. In fact, the
case of inference was one of the motivations for R. Solomonoff to introduce his
universal algorithmic probability \cite{solomonoff_formal_1964} as a tool to
reach an idealized inference machine, creating the notion of complexity concomitantly
to Kolmogorov. Subsequently, notions of complexity re-emerged in causal inference:
\cite{janzing_causal_2010} found, when a causal link exists between two random variables,
considering the direct joint probability is simpler, in terms of Kolmogorov complexity,
than the inverse direction.

More recently, \cite{marx_causal_2018} used Minimum Description Lenght to determine, given a joint probability distribution over $(X,Y)$, whether $X$ causes $Y$ or $Y$ causes $X$. Their method is based on tree models, and implying that a model respecting the causal relation will be simpler to describe.

\cite{tatti_finding_2008} PACK algorithm to cluster data using the simplest possible trees: maybe link this to isolation forests?

The topic of finding events from streams of time series data has already been explored in many ways.
\cite{aggarwal_outlier_2017} provides good review of modern approaches and techniques
in the field. Some previous work can also be noted for having used AIT techniques to qualify and detect events in time series data. For instance, \cite{batista_complexity-invariant_2011,fadlallah_weighted-permutation_2013} propose weighted permutation entropy as a proxy for complexity measures in time series data, and use it to find relations between different time series. \cite{hu_discovering_2011} proposes a MDL approach to find the intrinsic dimensions of time series. All these approaches are interesting, and can be integrated into our canvas as tools to detect events using only complexity. As such, one could acheive a purely complexity-driven process for detecting and qualifying events.

Among various methods, the approach of ``Isolation forests''\cite{liu_isolation_2008,hariri_extended_2021} is closely related, by design, to ours. It evaluates the isolation of data points by constructing random binary tree classifiers. On average, outlier points will require less operations to be singled out. Using the average height of leaves in the tree as a metrics, this approach succeeds in identifying outlier points without having to define a ``typical'' point. This approach can be understood in terms of complexity: each node of binary tree classifier needs a fixed amount of information to be described (which variable and threshold are used). So nodes that place higher need less information to be described. As such, outliers need less information to be singled out. Compared to ours, this method is tailored for data-points living in the same metric space. By using predicates as a proxy for complexity computation, our methods is more general, as it is agnostic regarding the nature of events. However, the introdcution of predicates adds a subjectivity in the determination of memorable points, as we will discuss later.

While all these works advocate in favor of a strong link between complexity and the
discovery of causes, not other works extends the notion up to the point we propose in this paper, using complexity only as a tool to express the intuitive notion of memorability, and using it for inference.

\section{Perspectives}
\label{sec:future}
The main purpose of the present paper is to show the possible connections
between existing definitions of simplicity from cognitive science, Algorithmic
Information Theory and a practical use case in cyber-physical systems. As such,
many further improvements can be done to pave the way towards a better
integration and performance for anomaly detection or abduction.

First, the main limitation of the current approach is the requirement of
predefined predicate concepts, from which the different filters are constructed.
As an extension, we suggest that in the future, we could explore online
generation of such predicate. A possibility would be to analyze discriminating
dimensions of incoming data and create predicate as to name these differences,
similar to the contrast operations proposed in \cite{dessalles_conceptual_2015,
  gardenfors2004conceptual}. For instance, the predicate concept $\mathtt{hot}$
can be discovered by discriminating a recent hot day along the
temperature axis and naming the difference with the prototypical day.

While the execution time is not part of the theoretical view of complexity, it
is of prime importance for practical applications, especially when one considers
implementation into real-time systems or embedded devices. While the computation
we propose appear to be heavy, and possibly heavier as the number of allowed
predicates grows, significant time savings can be achieve by trimming the base
memory of past events deemed the most ``non memorable''. For instance, one can
only retains the 100 most memorable events from the past. The difficulty with
this approach is that such operations should be done in a manner to not
interfere with the complexity computations for new elements: by forgetting some
past events, even uninteresting ones, one should make sure to keep track of what
made the interesting ones, interesting. Investigation of how to do so can pave
the way towards practical implementations and dynamic selection of interesting
events and help reducing the memory and computation cost of data-driven applications.

\section{Conclusion}

\bibliographystyle{IEEEtran}
\bibliography{biblio.bib}

\end{document}